---
title: "Report_Group_4_Up_All_Night-Rs"
author: "Tianyi Zuo, Roshael Chellappah, Tram Duong, Malcolm Stewart"
date: "2022/4/3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Soil microbial communities can play a key role in building plant communities. Also individual plants frequently cause changes in soil microbial community composition, and this reciprocal change then continues to affect the growth of neighboring plants or individual plants that subsequently colonize the soil. As this effect increases it can cause interspecific differences and thus affect plant community composition. In addition, the intensity of this interaction varies over time, so we focus on how the intensity of the interaction between plant and microbial populations varies over time. 

In our data, the experimenters mainly collected changes in the soil microbial community associated with individuals of four plant species more than 20 years. And the effect of soil biota on the performance of conspecific and heterospecific seedlings was quantified in a greenhouse experiment that preserved the soil characteristics of these monocultures. We used this dataset to investigate the associated effects of individual plants and soil microbes and whether they change in intensity over time. 
We propose relevant questions about this topic:

Question 1. How the Plant effect Bacterial community composition?

a)Does the plant species will affect the bacterial community composition?

b)Does the plant growth age will affect the bacterial community?

c)How do the bacterial communities composition change across time as the various plant species age?


Question 2. How the Bacterial community affect the plant species?

a). Do bacteria communities have an effect on plant root mass (below-ground dry mass)?

b). Which bacteria has the greatest effect on plant biomass when the plant is 4-12 years old?


## Package loading
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(ape)
library(vegan)
library(dplyr)
library(reshape2)
library(cowplot)
library(tidyverse)
library(ggpubr)
library(car)
```


## Input the data
```{r}
MBData<-read.csv("./Data/BacteriaCommunityMatrix.csv")
SBData<-read.csv("./Data/BacteriaSampleData.csv")
GData<-read.csv("./Data/GreenhouseData.csv")
```


## Question 1
Does the plant will affect the bacterial community in the soil source?


## Question 1-a
Does the plant species will affect the bacterial community composition?

In order to test the similarity between the bacterial community in different plant species area, we will use a Non-Metric Multidimensional Scaling (NMDS) to visualize differences among our samples within different plant species.  For this question, we mainly used the Bacteria Community Matrix to compared the composition of bacterial community among different plant species. The matrix data collected soil microbial communities associated with individuals of the four dominant plants species, which is community matrix of bacterial OTU characterized from each soil sample.

### Transforme the Data
According the data we have, I transform column and row first, and remove error which sum of row and column is zero. 
```{r}
rownames(MBData) <-MBData[,1] 
MBData<- MBData[,-1 ]
Matrix<-t(MBData)
Matrix <- Matrix[!(rowSums(Matrix) == 0), !(colSums(Matrix) == 0)]
```

### Distance Matrix
```{r warning=FALSE}
#Create the Bray_Curtis Dissimilarity metric
Dist<- vegdist(Matrix,method="bray",binary=F)
DistMat<-as.matrix(Dist)
PDat<-melt(DistMat)
```

### Non-Metric Multidimensional Scaling
Using Non-matric Dimensional Scaling(NMDS) to visualize the cluster of samplings from different plant species. We can find out whether in different plant species area has a stronger effect on bacterial communities.

```{r warning=FALSE}
set.seed(12138)
NMDSdat<-metaMDS(Dist,k=2  ,trymax = 200)
PDat<-data.frame(NMDS1=NMDSdat$points[,1],
                 NMDS2=NMDSdat$points[,2],
                 X=row.names(Matrix))
PDat<-merge(PDat,SBData,by="X",all.x=T,all.y=F)
qplot(x=NMDS1,y=NMDS2,colour=Species,alpha=I(0.6),data=PDat)+theme_bw()
```

Figure 1. NMDS plot show the overall variation in bacterial community composition among the different specific plant species. Points are color-coded based on the identity of the plant species: Ammophila (red); Baccharis(gree); Iceplant (orange); Lupine (blue); sand (pink).

### Discussion
According to the figure 1, the bacterial community composition differed between plant species, but the results were not significant. Based on the graph, the control group(sand) has less similarity with other four plant species.  Sampling from same plant specie's area are more likely cluster together, but there still have some overlap between different plant species. From figure 1, the bacterial community composition differed between plant species.


## Question 1-b
Does the plant growth age will affect the bacterial community composition?

We next wanted to investigate in more detail whether there is a temporal pattern in bacterial communities based on our available data and whether the composition of their communities changes depending on the time of plant growth.

### Transform Matrix
In order to test whether plant age will effect the bacterial community, we separate the bacterial community matrix into four different matrix depended on plant species. 
```{r}
MatrixA <- Matrix[grep("Ammophila", row.names(Matrix)), ]
MatrixB <- Matrix[grep("Baccharis", row.names(Matrix)), ]
MatrixI <- Matrix[grep("Iceplant", row.names(Matrix)), ]
MatrixL <- Matrix[grep("Lupine", row.names(Matrix)), ]
```

### Distance Matrix
```{r warning=FALSE}
#Create the Bray_Curtis Dissimilarity metric for each plant species
DistA<- vegdist(MatrixA,method="bray",binary=F)
DistMatA<-as.matrix(DistA)
PDatA<-melt(DistMatA)

DistB<- vegdist(MatrixB,method="bray",binary=F)
DistMatB<-as.matrix(DistB)
PDatB<-melt(DistMatB)

DistI<- vegdist(MatrixI,method="bray",binary=F)
DistMatI<-as.matrix(DistI)
PDatI<-melt(DistMatI)

DistL<- vegdist(MatrixL,method="bray",binary=F)
DistMatL<-as.matrix(DistL)
PDatL<-melt(DistMatL)
```

### Non-Metric Multidimensional Scaling
Using Non-matric Dimensional Scaling(NMDS) to visualize the cluster of samplings with plant age in four different plant species. We can find out whether the plant growth age has a stronger effect on bacterial communities.

```{r warning=FALSE}
#Non-Metric Multidimensional Scaling for Ammophila species.
set.seed(12138)
NMDSdatA<-metaMDS(DistA,k=2  ,trymax = 200)

PDatA<-data.frame(NMDS1=NMDSdatA$points[,1],
                 NMDS2=NMDSdatA$points[,2],
                 X=row.names(MatrixA))
PDatA<-merge(PDatA,SBData,by="X",all.x=T,all.y=F)
NMDSA<-qplot(x=NMDS1,y=NMDS2,colour=Age,alpha=I(0.6),data=PDatA) + theme_bw() + ggtitle("Ammophila")
```


```{r warning=FALSE}
#Non-Metric Multidimensional Scaling for Baccharis species.
set.seed(12138)
NMDSdatB<-metaMDS(DistB,k=2  ,trymax = 200)

PDatB<-data.frame(NMDS1=NMDSdatB$points[,1],
                 NMDS2=NMDSdatB$points[,2],
                 X=row.names(MatrixB))
PDatB<-merge(PDatB,SBData,by="X",all.x=T,all.y=F)
NMDSB<-qplot(x=NMDS1,y=NMDS2,colour=Age,alpha=I(0.6),data=PDatB)+theme_bw()+ ggtitle("Baccharis")
```

```{r warning=FALSE}
#Non-Metric Multidimensional Scaling for Iceplant species.
set.seed(12138)
NMDSdatI<-metaMDS(DistI,k=2  ,trymax = 200)

PDatI<-data.frame(NMDS1=NMDSdatI$points[,1],
                 NMDS2=NMDSdatI$points[,2],
                 X=row.names(MatrixI))
PDatI<-merge(PDatI,SBData,by="X",all.x=T,all.y=F)
NMDSI<-qplot(x=NMDS1,y=NMDS2,colour=Age,alpha=I(0.6),data=PDatI)+theme_bw() + ggtitle("Iceplant")
```

```{r}
#Non-Metric Multidimensional Scaling for Lupine species.
set.seed(12138)
NMDSdatL<-metaMDS(DistL,k=2  ,trymax = 200)

PDatL<-data.frame(NMDS1=NMDSdatL$points[,1],
                 NMDS2=NMDSdatL$points[,2],
                 X=row.names(MatrixL))
PDatL<-merge(PDatL,SBData,by="X",all.x=T,all.y=F)
NMDSL<-qplot(x=NMDS1,y=NMDS2,colour=Age,alpha=I(0.6),data=PDatL)+
  theme_bw()+
  ggtitle("Lupine")
```

```{r}
#Plotting  the NMDS result
plot_grid(NMDSA, NMDSB, NMDSI,NMDSL, ncol = 2)
```

Figure 2. The NMDS ordination plot shows the microbial community composition as factor of plant species and the age of plant individuals. Bacterial communities associated with the specific species are color-coded by individual plant age; the light to dark colour of point represent the age gradient from old to young, with species-specific minimum and maximum age.

### Discussion
According to figure 2, in each plant, the bacterial composition changes with the age of the plant. Among Ammophila species, the largest changes in bacterial community composition were observed. Bacterial abundance also changes with increasing plant age in the other three species but not significant. Overall, we can find the plant age is one of important factor which will cause the bacterial community change, but we still need furter research.  

## Question 1-c
How do the bacterial communities composition change across time as the various plant species age?

### Transform Data 
#### Subset Data - remove unwanted columns
```{r}
MDSubset <- GData %>% select(c("PlantSpecies", "SoilHostSpecies", "Age", "Total_Dry_mg")) # creates new dataset from MyData with only these 4 columns
```

#### Data Prepation
The purpose of this section is to identify whether the bacterial soil host species composition is in any way correlated with or impacted by the age of the examined plant species. From answering this question and if some signification correlation is identified, we can further look into why bacterial soil host species compositions may differ across time. 

First, we will first filter the data by plant species:
```{r}
AaSub <- MDSubset %>% filter(PlantSpecies == "Aa") # creates subset of MDSubset with only Aa plant data
BpSub <- MDSubset %>% filter(PlantSpecies == "Bp")
CeSub <- MDSubset %>% filter(PlantSpecies == "Ce")
LaSub <- MDSubset %>% filter(PlantSpecies == "La")
```

Since both age and bacterial soil host species are categorical variables, it is difficult to statistically identify whether the relationship between these to variables is significant or not. Therefore, we believe that the best way to identify any difference in bacterial soil host species composition across time would be to visualize the data. Therefore, we will create stacked bar graphs for each plant species, identifying the approximate bacterial compositions within plant species at every age that samples were collected. 

```{r}
# Graph for Aa
AaPlot <- ggplot(data = AaSub, aes(x = Age, fill = SoilHostSpecies)) +
  geom_bar() +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) + 
  scale_y_continuous(breaks = seq(0, 20, by = 2)) +
  ggtitle("Ammophila arenaria") +
  xlab("Age (years)") + ylab("Number of Samples") +
  guides(fill=guide_legend(title="Soil Host Species")) +
  theme_bw()

# Graph for Bp
BpPlot <- ggplot(data = BpSub, aes(x = Age, fill = SoilHostSpecies)) +
  geom_bar() +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +  
  scale_y_continuous(breaks = seq(0, 20, by = 2)) +
  ggtitle("Baccharis pilularis") +
  xlab("Age (years)") + ylab("Number of Samples") +
  guides(fill=guide_legend(title="Soil Host Species")) +
  theme_bw()

# Graph for Ce
CePlot <- ggplot(data = CeSub, aes(x = Age, fill = SoilHostSpecies)) +
  geom_bar() +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +  
  scale_y_continuous(breaks = seq(0, 20, by = 2)) +
  ggtitle("Carpobrotus edulis") +
  xlab("Age (years)") + ylab("Number of Samples") +
  guides(fill=guide_legend(title="Soil Host Species")) +
  theme_bw()

# Graph for La
LaPlot <- ggplot(data = LaSub, aes(x = Age, fill = SoilHostSpecies)) +
  geom_bar() +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +  
  scale_y_continuous(breaks = seq(0, 20, by = 2)) +
  ggtitle("Lupine arboreus") +
  xlab("Age (years)") + ylab("Number of Samples") +
  guides(fill=guide_legend(title="Soil Host Species")) +
  theme_bw()

# Arrange all these graphs together on one page
Plot <- ggarrange(AaPlot, BpPlot + rremove("ylab"), CePlot, LaPlot + rremove("ylab"), # helps identify which plots to use within the multiplot figure, and rremove("ylab") removes the y-axis label for 2 of the 4 graphs
          labels = c("A", "B", "C", "D"), # sets labels for every plot
          ncol = 2, nrow = 2, # sets the dimensions for the multiplot
          common.legend = TRUE, legend = "right") # sets a common plot and places it on the left of the page

annotate_figure(Plot,
                top = text_grob("Bacterial Soil Host Species Compositions Per Plant Species", color = "black", face = "bold", size = 14)) # creates a common title for the multiplot
```

Figure 3. Stacked bar graphs of bacterial soil host species for every plant species sampled. Y-axis identifies the numbers of samples collected at each age group and bacterial composition can be identified by the various colours within every bar. Greater amount of one particular colour within one bar indicates a greater abundance of that bacteria at that specific age. 

### Discussion
One conclusion that can be made from these stacked bar graphs is that the bacterial species Lupine disappears from the soil environments of all 4 plant species after 13 years of age. Although there were less total *Baccharis pilularis* samples taken in the Greenhouse experiment, all plant species began with a similar composition of all studied bacterial soil host species. Furthermore, it also seems that *Ammophila arenaria* is always in greatest abundance when the plants are 26 years old. All plant species also seem to follow similar trends in how the bacterial community composition fluctuates over time. Further research could identify why Lupine suddenly disappears from the soil microbiome after 13 years of age, and if there is some other factor that results in this disappearance. It would also be useful to identify whether the relationship between age and bacterial soil host species is significant; however, since both of these variables are categorical and since the data is not normally distributed, it is difficult to identify a particular statistical test that could help us identify this significance. 



## Question 2-a 
Do bacteria species have an effect on plant root mass (below-ground dry mass)?  

In order to account for different root masses among species of plants, the relationship between soil host species and root mass will be looked at per plant species. In order to account for difference in root masses among age, an older age group and younger sampled age group will be chosen among all plant species (10 years and 0.5 years, respectively) for visual analysis.
In order to account for differences in root mass by age and for statistical confirmation, the ANCOVA test will be used.

### Transform the Data
```{r}
#Subset the data to get only the columns of interest
root <- GData[c("PlantSpecies", "SoilHostSpecies", "BG_Dry_mg", "Age")]
 
#Age needs to be changed into a categorical variable
class(root$Age)
root$Age <- as.factor(root$Age)
class(root$Age)
```

### Visualizing the data:
```{r}

#Subset data again to get only the ages of interest
 oldplants <- root[grep('10', root$Age),]
 youngplants <- root[grep('0.5', root$Age),]
```
```{r}
oldplot <- ggplot(oldplants, aes(x=PlantSpecies, y=BG_Dry_mg, fill=SoilHostSpecies)) + 
    geom_boxplot() +
    facet_wrap(~PlantSpecies, scale="free") +
  labs(x= "Plant species (15 years)", y = "Below-ground dry mass (mg)", fill= "Soil host species") +
  theme_classic()
oldplot
```

Figure 4. Box plot of the relationship between soil host species and below-ground dry mass in milligrams of each plant species at age 10 years (A. arenaria, B. pilularis, C. edulis, L. arboreus, respectively). Bp at age 15 was not treated with Lupine for an unknown reason (possibly due to lack on samples).

```{r}
youngplot <- ggplot(youngplants, aes(x=PlantSpecies, y=BG_Dry_mg, fill=SoilHostSpecies)) + 
    geom_boxplot() +
    facet_wrap(~PlantSpecies, scale="free") +
  labs(x= "Plant species (0.5 years)", y = "Below-ground dry mass (mg)", fill= "Soil host species") +
  theme_classic()
youngplot
```

Figure 5. Box plot of the relationship between soil host species and below-ground dry mass in milligrams of each plant species at age 0.5 years (A. arenaria, B. pilularis, C. edulis, L. arboreus, respectively).


### Statistical analysis using ANCOVA:
```{r}
#Checking data to meet assumptions

#Assumption 1: the covariate and treatment should be independent of each other.
root$Age <- as.numeric(root$Age)
rootmodel <- aov(Age ~ SoilHostSpecies, data = root)
summary(rootmodel)
```

```{r warning=FALSE}
#Assumption 2: There is homogeneity of variance.
leveneTest(BG_Dry_mg~SoilHostSpecies, data = root)
```

The covariance model ANCOVA will be used to statistically analyse for significance in soil host species.

```{r}
#Take age into consideration
ancova_model <- aov(BG_Dry_mg ~ SoilHostSpecies + PlantSpecies + Age, data = root)
Anova(ancova_model, type="III")
```

### Discussion
In Figures 4 and 5, box plots mostly overlap with each other, showing weak correlation between soil host species and below-ground dry mass. Across species, there doesn't appear to be a prominent soil host species contributing to below-ground dry mass. Across ages 0.5 to 10 years for each species, there doesn't appear to be a prominent soil host species contributing to below-ground dry mass.
According to the ANCOVA analysis, the p-value for soil host species is more than 0.05. This indicates that soil host species are not significantly correlated to below-ground dry mass, even when taking age into account.
Therefore, it is assumed that the bacteria community formed by the soil host species is not significant to plant root growth.


## Question 2-b
Which bacteria has the greatest effect on plant biomass when the plant is 4-12 years old?

The goal of this section is to determine which bacteria has the greatest effect on plant biomass when the plant is between 4 and 12 years old. This was chosen as this is the average age of the plants in the data set.

### Transform the Data 
```{r}
# Removing all but wanted data points
GNdata = GData %>%
  select(BG_Dry_mg, PlantSpecies, SoilHostSpecies, Age)
```

#### Setting up Data

```{r}
#create a new data set with only the plants that are this old
subdata = GNdata %>%
  subset(Age == 4:12)
head(subdata)
```

Now with this data I can begin my analysis

### Analysis

I will use a box plot to see the impact that the bacteria have on the plants that are this age
```{r}
Plot1 = ggplot(subdata, aes(x = SoilHostSpecies, y = BG_Dry_mg, fill = SoilHostSpecies)) +
  geom_boxplot() +
  theme_classic() +
    scale_fill_discrete(name = "Soil Bacteria") +
    scale_x_discrete(name = "", labels = c("","","","")) +
    scale_y_continuous(name = "Root Dry Mass (mg)")
Plot1
```
Figure 6: Shows the relationship between the different species of Soil Bacteria and Root Dry mass in mg. Lupine has the highest average value while IcePlant has the greatest variation



```{r}
new_subdata = subdata
  for (i in 1:length(subdata$PlantSpecies)) {
    if (subdata$SoilHostSpecies[i] != "Lupine"){
      new_subdata$SoilHostSpecies[i] = "Other"
    }
  }
```

Now I will recreate the plot with the new data set
```{r}
Plot2 = ggplot(new_subdata, aes(x = SoilHostSpecies, y = BG_Dry_mg, fill = SoilHostSpecies)) +
  geom_boxplot() +
  theme_classic() +
    scale_fill_discrete(name = "Soil Bacteria") +
    scale_x_discrete(name = "", labels = c("","")) +
    scale_y_continuous(name = "Root Dry Mass (mg)")
Plot2
```
Figure 7: Shows the difference in Root Dry Mass between Lupine and all the other different species of soil bacteria.


Now I will conduct an unpaired two-sample t-test to determine if this difference between the two groups is significant or not. I am using a two-sample t-test as the two groups are unrelated.

```{r}
t.test(subset(new_subdata$BG_Dry_mg, new_subdata$SoilHostSpecies == "Other"),
       subset(new_subdata$BG_Dry_mg, new_subdata$SoilHostSpecies == "Lupine"),
              alternative = "two.sided", var.equal = FALSE)
```

Table 1: Shows the results of a Welch Two Sample t-test of below ground dry mass of plants with Lupine against plants with any of the other different species. With a p-value of 0.6816, the results are not statistically significant

### Discussion
According to figure 6, we can see that Lupine has an impact on the dry weight of the roots between 4 and 12 years. To further examine this relationship, I will create a new set of data that groups together all the other species of soil bacteria, then create a boxplot using that set.
Based on graph 7 we can see visually that there is a difference between samples with Lupine and all the other species of soil bacteria at this point in time. 
Besides, based on the results of this two-sample t-test, we can see that while there is a difference between the two groups, this difference is not enough for it to be considered significant as the p-value is not less that 0.05.


## Conclusion

Our data suggest that plant-soil microbial interactions do exist, but we were not able to demonstrate whether their strength is affected by the age of the individual plant. In our data, the different ages of plant individuals can differ but not significantly on bacterial community composition. The same result was found in the analysis of soil microorganisms affecting plant root weight. At present, it seems that the age of the individual plant does not affect the interaction between the plant and the soil microorganisms. That is different from the results in our data source article. It is suggested in the article that the strength of this interaction correlates with the length of time that individual plants condition the soil. That may be because other influencing factors, such as soil conditions, were not included in our analysis, thus affecting the different results. However, the difference of microbial samples due to plant age was also found in our data, suggesting that individual plant age may be a potential factor that could influence the interaction between plant and soil microbes. Subsequent studies could focus more on whether the length of soil conditioning of plants affects the strength of this interaction and should allow for more appropriate placement of experimental results in natural settings to predict better how soil microbes affect plant community assembly.


